### Author: Brandon Rodriguez-Hernandez

### Published: January 27, 2025

**Abstract:**

This is a technical blog post of **both** an HTML file *and* [.qmd file](wine_features.qmd) hosted on GitHub pages.

**Step Up Code:**

```{r}
library(tidyverse)
library(caret)
library(fastDummies)
wine <- readRDS(gzcon(url("https://github.com/cd-public/D505/raw/master/dat/wine.rds")))
```

**Explanataion:**

> [TODO]{style="color:red;font-weight:bold"}: *The first three lines pull in the libraries needed, and the line after that downloads wine.rds into the qmd file.*

# Feature Engineering

We begin by engineering an number of features.

1.  Create a total of 10 features (including points).
2.  Remove all rows with a missing value.
3.  Ensure only log(price) and engineering features are the only columns that remain in the `wino` dataframe.

```{r}
wino <- wine %>% 
  mutate(lprice=log(price),
         
      
          points_category = case_when(
            points >= 90 ~ "Excellent",
            points >= 85 ~ "Good",
            TRUE ~ "Fair"
          ),
      
          variety_complexity = str_count(variety, "\\s+") + 1,
      
          #has_designation = ifelse(!is.na(designation), "Yes", "No"),
         
          description_length = str_count(description, "\\s+") + 1,
          
          points_deviation = points - ave(points, variety, FUN = mean),

          



          )%>%
  tidyr::drop_na() %>%  # Use tidyr::drop_na() for clarity. dplyr also has drop_na but tidyr:: is best practice.
  # Select only the engineered features (and lprice, if that's your target variable)
  select(lprice, points_category, variety_complexity, 
          description_length, points_deviation)



```

# Caret

We now use a train/test split to evaluate the features.

1.  Use the Caret library to partition the wino dataframe into an 80/20 split.
2.  Run a linear regression with bootstrap resampling.
3.  Report RMSE on the test partition of the data.

```{r}
# TODO: hint: Check the slides.
wine_index <- createDataPartition(wino$lprice, p = 0.8, list = FALSE)
wino_tr <- wino[wine_index, ]
wino_te <- wino[-wine_index, ]

ctrl <- trainControl(method = "boot", 
                    number = 100,     
                    savePredictions = TRUE, 
                    verboseIter = FALSE)

# Train the linear model with bootstrapping
m1_boot <- train(lprice ~ .,
                data = wino_tr,
                method = "lm",
                trControl = ctrl)


predictions_boot <- predict(m1_boot, newdata = wino_te)


rmse_test <- RMSE(predictions_boot, wino_te$lprice)

rmse_test


```

# Variable selection

We now graph the importance of your 10 features.

```{r}
# TODO: hint: Check the slides.

importance <- varImp(m1_boot)


importance_df <- importance$importance
importance_df$Feature <- rownames(importance_df) # Add feature names as a column

ggplot(importance_df, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(fill = "skyblue", color = "black") +
  coord_flip() +  # Horizontal bars
  labs(title = "Variable Importance", x = "Feature", y = "Importance") +
  theme_bw() # Clean theme




```
